Estimating duration in the presence of misclassification

### 2024-03-19 conclusion.tex

In \cref{E-transmission}, I found that this it is possible, developing two models that can infer incidence and transmission from prevalence surveys.
One model was phenomenological in nature and the other mechanistic.
The phenomenological approach, based on the backcalculation framework, provide estimates of incidence and makes fewer assumptions about the infection process; however, the estimates are very uncertain.
The mechanistic approach, based on a SIR-type compartmental model, reduces the uncertainty and allows estimation of the components of transmission by making stronger assumptions about the infection process.
These results highlighted the heterogeneity across regions and age groups in England.

@@

These models both required knowledge of the distribution of the duration of positivity.
Estimating this distribution was the focus of the majority of this thesis.
In \cref{E-perf-test,E-imperf-test}, I showed how this quantity could be estimated the CIS prevalence survey.
This did require external information from the ATACCC study (the results of \cref{E-ATACCC}), demonstrating that the CIS's design is not sufficient for estimating transmission.
However, future surveys may be able to avoid this issue (see \cref{conclusion:sec:future-work}).

@@

Overall, I have found that representative prevalence surveys are a promising approach for pandemic surveillance.
A well-designed survey is sufficient to estimate all of the quantities of interest (as described in \cref{E-intro:sec:metrics}), except severity.

@@

The feasibility of jointly modelling duration and incidence was not investigated in this thesis.
The model for duration in \cref{E-perf-test,E-imperf-test} allowed estimation of the total number of infections in the cohort but assumed these occurred uniformly over the duration of the survey.
If, rather than assuming uniformity, the time of incidence was estimated, this would be a form of incidence estimation.
The issue of including covariates to control for non-response bias, encountered in \cref{E-transmission}, as well as the parameterisation of the shape of the incidence curve would be major challenges in this extension.
The basis of such a model could be existing modelling frameworks~\autocite[e.g.][]{taffeJoint,haySerosolver}.
These frameworks addressed the challenges by: taking approximations (\eg empirical Bayes), having smaller sample sizes, or using high-performance computing.
However, the CIS is larger and incidence estimation desired at a more granular level than previously encountered.
Furthermore, high-performance computing cannot be used within the SRS.

@@

This thesis did not consider detection or early surveillance of emerging pandemics.
Initial uncertainty about local transmission in the UK may have led to suboptimal decisions at the start of 2020~\autocite{pellisChallenges}.
Community testing was not widely available at this stage of the pandemic, limiting the knowledge of the extent of local (as opposed to imported) transmission~\autocite{whittyTechReportCOVID}.
A representative survey of the population could have provided this information.
However, challenges in constructing a survey with the power to detect low levels of transmission, as well as the time required to set-up such a survey, would need to be addressed.

A final consideration is the estimation of the severity of a pathogen.
Severity of a pathogen is crucial for risk assessment, and can change over time~\autocite[e.g.][]{kirwanSeverityTrends}.
Two approaches could be taken to using a prevalence surveys for this purpose.
First, a cohort approach where the outcomes of individuals who test positive are followed up.
The low number of positive cases in the CIS would make this challenging.
In addition, the data would need to be linked to relevant records.
Second, a transmission model, such as the one in \cref{E-SEIR}, could jointly model severe events and prevalence.
Such an approach has previously been used~\autocite{daviesAssociation,ironsEstimating,knockKey,nicholsonImproving,pooleyEstimation,birrellRTM2} but without the sophistication of the observation model developed in \cref{E-SEIR:sec:observation}.

>>>>

In \cref{E-transmission}, I found that this it is possible, developing two models that can infer incidence and transmission from prevalence surveys.
One model was phenomenological in nature and the other mechanistic.
The phenomenological approach, based on the backcalculation framework, provides estimates of incidence and makes fewer assumptions about the infection process; however, the estimates are very uncertain.
The mechanistic approach, based on an SIR-type compartmental model, reduces the uncertainty and allows estimation of the components of transmission by making stronger assumptions about the infection process.
These results highlighted the heterogeneity across regions and age groups in England.

@@

These models both required knowledge of the distribution of the duration of positivity.
Estimating this distribution was the focus of most of this thesis.
In \cref{E-perf-test,E-imperf-test}, I showed how this quantity could be estimated the CIS prevalence survey.
This did require external information from the ATACCC study (the results of \cref{E-ATACCC}), demonstrating that the CIS's design is not sufficient for estimating transmission.
However, future surveys may be able to avoid this issue (see \cref{conclusion:sec:future-work}).

@@

Overall, I have found that representative prevalence surveys are a promising approach for pandemic surveillance.
A well-designed survey is sufficient to estimate all the quantities of interest (as described in \cref{E-intro:sec:metrics}), except severity.

@@

The feasibility of jointly modelling duration and incidence was not investigated in this thesis.
The model for duration in \cref{E-perf-test,E-imperf-test} allowed estimation of the total number of infections in the cohort but assumed these occurred uniformly over the duration of the survey.
If, rather than assuming uniformity, the time of incidence was estimated, this would be a form of incidence estimation.
The issue of including covariates to control for non-response bias, encountered in \cref{E-transmission}, as well as the parameterization of the shape of the incidence curve would be major challenges in this extension.
The basis of such a model could be existing modelling frameworks~\autocite[e.g.][]{taffeJoint,haySerosolver}.
These frameworks addressed the challenges by: taking approximations (\eg empirical Bayes), having smaller sample sizes, or using high-performance computing.
However, the CIS is larger and incidence estimation desired at a more granular level than previously encountered.
Furthermore, high-performance computing cannot be used within the SRS.

@@

This thesis did not consider detection or early surveillance of emerging pandemics.
Initial uncertainty about local transmission in the UK may have led to suboptimal decisions at the start of 2020~\autocite{pellisChallenges}.
Community testing was not widely available at this stage of the pandemic, limiting the knowledge of the extent of local (as opposed to imported) transmission~\autocite{whittyTechReportCOVID}.
A representative survey of the population could have provided this information.
However, challenges in constructing a survey with the power to detect low levels of transmission, as well as the time required to set up such a survey, would need to be addressed.

A final consideration is the estimation of the severity of a pathogen.
Severity of a pathogen is crucial for risk assessment and can change over time~\autocite[e.g.][]{kirwanSeverityTrends}.
Two approaches could be taken to using a prevalence surveys for this purpose.
First, a cohort approach where the outcomes of individuals who test positive are followed up.
The low number of positive cases in the CIS would make this challenging.
In addition, the data would need to be linked to relevant records.
Second, a transmission model, such as the one in \cref{E-SEIR}, could jointly model severe events and prevalence.
Such an approach has previously been used~\autocite{daviesAssociation,ironsEstimating,knockKey,nicholsonImproving,pooleyEstimation,birrellRTM2} but without the sophistication of the observation model developed in \cref{E-SEIR:sec:observation}.

### 2024-03-19 introduction-comments.tex
remove appendix from chapter-by-chapter compile

\Textcite{mccabeCISincidence}, adapting the methodology \textcite{ealesAppropriately} developed for REACT, estimate $\Re$ and the growth rate from the CIS data.
They estimate a posterior distribution of the parameters of a spline by fitting to the prevalence results via a binomial distribution.
Their setup penalises the spline deviating from a constant growth rate.
The growth rate is then calculated from the derivative of the spline, an analytical transformation of the parameters.
Since, under reasonable model assumptions, there is a one-to-one correspondence between reproduction numbers and growth rates~\autocite{wallingaGI}, and they assume that the growth rate in incidence and prevalence is equal, this also estimates $\Re$.

@@

Modelling the duration directly, as I do in \cref{E-perf-test,E-imperf-test} considers only the binary result at each test (positive or negative); in this thesis a \emph{survival analysis} framework is adopted.
This approach generally requires fewer parameters, and no random effect(s), reducing the computational burden.
It is also easier to specify non-parameteric forms of the distribution, which is important for SARS-CoV-2 as the tail behaviour may be important.
A more detailed review of the literature is in the relevant chapters.
% Survival analysis is the area of statistics concerned with estimating the distribution of the times between two events.
Other related approaches, such as a multi-state model (reviewed by \textcite{jacksonMSM}), could also be considered; multi-state models can be viewed as a generalisation of survival analysis.

@@

\ifSubfilesClassLoaded{
  \appendix
  \subfile{MCMC-appendix}
  \listoftodos
}{}
\end{document}

>>>>

\Textcite{mccabeCISincidence}, adapting the methodology \textcite{ealesAppropriately} developed for REACT, estimate $\Re$ and the growth rate from the CIS data.
They estimate a posterior distribution of the parameters of a spline by fitting to the prevalence results via a binomial distribution.
Their setup penalizes the spline deviating from a constant growth rate.
The growth rate is then calculated from the derivative of the spline, an analytical transformation of the parameters.
Since, under reasonable model assumptions, there is a one-to-one correspondence between reproduction numbers and growth rates~\autocite{wallingaGI}, and they assume that the growth rate in incidence and prevalence is equal, this also estimates $\Re$.

@@

Modelling the duration directly, as I do in \cref{E-perf-test,E-imperf-test} considers only the binary result at each test (positive or negative); in this thesis a \emph{survival analysis} framework is adopted.
This approach generally requires fewer parameters, and no random effect(s), reducing the computational burden.
It is also easier to specify non-parametric forms of the distribution, which is important for SARS-CoV-2 as the tail behaviour may be important.
A more detailed review of the literature is in the relevant chapters.
% Survival analysis is the area of statistics concerned with estimating the distribution of the times between two events.
Other related approaches, such as a multi-state model (reviewed by \textcite{jacksonMSM}), could also be considered; multi-state models can be viewed as a generalisation of survival analysis.

@@

\ifSubfilesClassLoaded{
  \appendix
%   \subfile{MCMC-appendix}
  \listoftodos
}{}
\end{document}

### 2024-03-19 introduction.tex

\emph{Transmission} refers to the process by which an infectious disease generates new infections.
Estimating transmission refers to inferring the parameters of this process, including its components.
The components of transmission (\eg the number of contacts each individual makes per day) allow interventions to be targeted.
They are also important for forecasting future incidence.

@@

The data sources used for outbreak surveillance are often observational in nature.
Therefore, they suffer from biases that can be hard to understand and quantify.
Here I discuss three types of data which are useful for surveillance.
Incidence-type and prevalence-type data can directly inform estimation of incidence and transmission; however, this estimation almost always relies on prior knowledge of epidemiological and/or biological quantities.
These quantities can be estimated from other studies, \eg cohort studies.

@@

Next, are semi-mechanistic models.
These have components of the transmission process in them but do not fully represent the process.
The most common type of semi-mechanistic model is those based on a \emph{renewal equation}.
The renewal equation~\autocite{fraserRenewal} for an infectious disease, in the most general form, is:
\begin{align}
    Z(t) &= \int_{0}^{t} \beta(t,\tau) Z(t-\tau) d \tau
\end{align}
where $Z(t)$ is the mean incidence at time $t$ and $\beta(t, \tau)$ is the instantaneous probability of an individual infected $\tau$ time units ago infects an individual at time $t$.
Therefore, the renewal equation relates incidence in the past to incidence in the future.
The effective reproduction number can be expressed as a function of $\beta(t, \tau)$.
% A scaled measure of $I(t)$ (\eg hospital admissions or community test results) allows estimating $\beta$.
% The most common implementations are the R packages EpiNow2~\autocite{EpiNow2} and EpiEstim~\autocite{EpiEstim}.
Renewal equations do not feature further in this thesis; for details, including practical considerations when using them as the basis for inference, see \textcite{nashEstimating,gosticPractical}.

Finally, mechanistic models are the most complex, representing the transmission process in varying levels of detail.
\emph{Compartmental} models, the most common type of mechanistic model, split the population into compartments, each representing a different state of the disease.
The \emph{dynamics} of the epidemic, how it changes over time, are represented by the transitions between compartments (for more details see \cref{E-SEIR:sec:mechanistic-models}).
Simple mechanistic models use a homogeneous population, with individuals differing only due to their disease state, and exponential distributions for the time in each disease state.
More complex models relax these assumptions, in particular incorporating heterogeneity; the age structure of the population is important for respiratory diseases.
% Mathematically, the dynamics are described by ODEs (ordinary differential equations) or SDEs (stochastic differential equations).
% Unlike renewal equations, there are not widely-used implementations of compartmental models.
% A commonly used frameork is the SEIR (susceptible, exposed, infectious, recovered) model.
% The SEIR model, has four components (susceptible, exposed, infectious, recovered), with the population divided into these components; it is described in detail in \cref{E-SEIR:sec:mechanistic-models}.

@@

The \emph{disease model} links the transmission model to a quantity, $W_t$, observable albeit possibly in a noisy and/or biased way (see next section).
Examples are the incidence or prevalence of symptomatic individuals and admissions to hospital.

The choice of observation model can, in theory, be independent of the transmission model.
However, pragmatic considerations often mean choosing the observation model to be in a similar form to the transmission model.

Typically, the core of an observation model is a \emph{convolution equation} or an approximation of it.
This equation describes the expectation of $W_t$, conditional on the incidence prior to time $t$.
In discrete time, it is:
\begin{align}
    \E(W_t \mid Z_1, Z_2, \dots) &= \zeta \sum_{i=0}^\infty Z_{t-i} f(i)
    \label{intro:eq:disease-model}
\end{align}
where $Z_t$ is the incidence at time $t$; $\zeta$ is the proportion of infections that ever appear in $W_t$; and $f(i)$ is the probability that an infection at time $t-i$ is included in count $W_t$, conditional on it ever being in one of the counts.
$f$ is known as a \emph{delay distribution}.

>>>>

\emph{Transmission} refers to the process by which an infectious disease generates new infections.
Estimating transmission refers to inferring the parameters of this process.
The components of transmission (\eg the number of contacts each individual makes per day) allow interventions to be targeted.
They are also important for forecasting future incidence.

@@

The data sources used for outbreak surveillance are often observational in nature.
Therefore, they suffer from biases that can be hard to understand and quantify.
Here I discuss three types of data which are useful for surveillance.
Incidence-type and prevalence-type data can directly inform estimation of incidence and transmission; however, this estimation almost always relies on prior knowledge of epidemiological and/or biological quantities.
For example, the time between infection and symptom onset.
These quantities can be estimated from other studies, \eg cohort studies.

@@

Next, are semi-mechanistic models.
These have components of the transmission process in them but do not fully represent the process.
The most common type of semi-mechanistic model is those based on a \emph{renewal equation}.
The renewal equation~\autocite{fraserRenewal} for an infectious disease, in the most general form, is:
\begin{align}
    Z(t) &= \int_{0}^{t} \beta(t,\tau) Z(t-\tau) d \tau
\end{align}
where $Z(t)$ is the mean incidence at time $t$ and $\beta(t, \tau)$ is the mean rate an individual infected $\tau$ time units ago infects individuals at time $t$.
Therefore, the renewal equation relates incidence in the past to incidence in the future.
This partially describes the transmission process, hence the name semi-mechanistic.
The effective reproduction number can be expressed as a function of $\beta(t, \tau)$.
% A scaled measure of $I(t)$ (\eg hospital admissions or community test results) allows estimating $\beta$.
% The most common implementations are the R packages EpiNow2~\autocite{EpiNow2} and EpiEstim~\autocite{EpiEstim}.
Renewal equations do not feature further in this thesis; for details, including practical considerations when using them as the basis for inference, see \textcite{nashEstimating,gosticPractical}.

Finally, mechanistic models are the most complex, representing the transmission process in varying levels of detail.
\emph{Compartmental} models, the most common type of mechanistic model, split the population into compartments, each representing a different state of the disease.
The \emph{dynamics} of the epidemic, how it changes over time, are represented by the transitions between compartments (for more details see \cref{E-SEIR:sec:mechanistic-models}).
Simple mechanistic models use a homogeneous population, with individuals differing only due to their disease state, and exponential distributions for the time in each disease state.
More complex models relax these assumptions, in particular to incorporate heterogeneity.
The age structure of the population is important for respiratory diseases because individuals meet individuals of their own age more frequently.
% Mathematically, the dynamics are described by ODEs (ordinary differential equations) or SDEs (stochastic differential equations).
% Unlike renewal equations, there are not widely-used implementations of compartmental models.
% A commonly used frameork is the SEIR (susceptible, exposed, infectious, recovered) model.
% The SEIR model, has four components (susceptible, exposed, infectious, recovered), with the population divided into these components; it is described in detail in \cref{E-SEIR:sec:mechanistic-models}.

@@

\emph{Disease models} link the transmission model to a latent quantity, $W_t$.
$W_t$ will then be recorded through some system (\eg a testing programme) described by the observation model (see next section).
For example, $W_t$ could be the incidence or prevalence of symptomatic individuals on day $t$.

At the core of many observation models is a \emph{convolution equation} or an approximation of it.
This equation describes the expectation of $W_t$, conditional on the incidence prior to time $t$.
In discrete time, it is:
\begin{align}
    \E(W_t \mid Z_1, Z_2, \dots) &= \zeta \sum_{i=0}^\infty Z_{t-i} f(i)
    \label{intro:eq:disease-model}
\end{align}
where $Z_t$ is the incidence at time $t$; $\zeta$ is the proportion of infections that ever appear in $W_t$; and $f(i)$ is the probability that an infection at time $t-i$ is included in count $W_t$, conditional on it ever being in one of the counts.
$f$ is known as a \emph{delay distribution}.

### 2024-03-19 transmission.tex

First, in \cref{backcalc}, I use a back-calculation approach, based on \cref{E-inc-prev:eq:EPt}.
This approach enables estimation of incidence within (sub)populations of England.
There are several attractive features about this approach in the context here.
The backcalculation approach makes no assumptions about the transmission process; the assumptions have the potential to introduce bias.
It can also easily incorporate arbitrary covariates to adjust for non-response bias.
The scale of the CIS and desire to correct for non-response bias presents challenges for this methodology, especially within the limited computation of the SRS environment.

@@

\subsubsection{Possible approaches}

The first approach, which I call the \emph{incidence-inferred} approach, is a single statistical model where incidence is inferred, with prevalence as a derived quantity.
The basis of the statistical model, in a Bayesian framework, would be as follows.
\begin{align}
    \vec{Z} \mid \vec{\phi} &\dist g_Z(\vec{\phi}) \\
    P_{i,t} &= \matr{S} \vec{Z}_i \\
    Y_{i,t} \mid \vec{P} &\dist \BB(n_{i,t}, P_{i,t})
\end{align}
where $\vec{Z}$ is a vector of incidence proportion for each strata and day defined by some model $g$ with parameters $\vec{\phi}$; $P_{i,t}$ is the prevalence in stratum $i$ on day $t$; $\matr{S}$ is the survival matrix defined in \cref{E-inc-prev:eq:Smatrix}; $Y_{i,t}$ is the number of positive tests in stratum $i$ on day $t$; and $n_{i,t}$ is the number of tests in stratum $i$ on day $t$.
The model $g_Z$ would be specified to share information across time and/or strata.
This both helps stabilise the estimation (see \cref{intro:sec:disease-model}) and allows borrowing of information across strata.
Borrowing information is required because some strata and time combinations have few positives (fewer than five per day).
The relationship between $P$ and $Z$ is from \cref{E-inc-prev:eq:EPt-matrix}, although dividing both sides by $\Npop$ to have the values in terms of incidence proportion and prevalence rather than incidence and number of prevalent individuals.

The second approach, which I call the \emph{prevalence-inferred} approach takes a two-step approach.
First, it uses the statistical model from \cref{E-biology-data:sec:MRP} to estimate prevalence.
Then, it transforms the posterior distribution of prevalence to a posterior distribution of incidence using the forward substitution as in \cref{E-inc-prev:eq:forward-substitute}.
At a high-level, the statistical model is:
\begin{align}
    \vec{P} \sim \vec{\phi} &\dist g_P(\vec{\phi}) \\
    Y_{i,t} &\dist \BB(n_{i,t}, P_{i,t})
\end{align}
where $g_P$ is analogous to $g_Z$ in providing a joint prior for the prevalence in each stratum and day combination.

\subsubsection{Desirable properties}

The incidence-inferred approach is more parsimonious in being a single step procedure.
However, it poses several pragmatic challenges which means I favour the prevalence-inferred approach.
Here I explain that further developments, beyond the scope of this thesis, would be required to use the incidence-inferred approach with a suitable $\pi_Z$ within the computational constraints of the SRS.

@@

The chosen approach should have the following properties.
\begin{enumerate}
    \item The smoothing model over time should penalize changes from continuous exponential growth.
        penalizing changes from continuous exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalization has previously been shown to improve estimates of SARS-CoV-2 prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}.
    \item Limit estimates to epidemiologically and biologically plausible values.
        That is, incidence proportion and prevalence should be in the interval $[0, 1]$.
        % While incidence proportion can theoretically be greater than 1, this would imply that some individuals were infected multiple times in a day which is not biologically plausible.
    \item Have sufficient parameters to incorporate all necessary covariates to correct for non-response bias.
    \item Borrow information across the different covariates.
        The borrowing should penalize changes in the distribution of incidence/prevalence across strata in the same geographic region being constant.
        A constant distribution is predicted by the theoretical study of epidemics, for example the dynamics of stratified, mechanistic transmission models (such as the one introduced in \cref{E-SEIR:sec:structured-populations}) over short periods of time with negligible behavioural and immunity changes.
    \item Be computationally efficient.
        The model will be fit to granular data and therefore needs to be run within the SRS, with its computational limitations (see \cref{E-biology-data:sec:SRS}).
\end{enumerate}

Properties 1 and 2 suggest a model of the form $\logit(\pi_{i,t}) = g(i, t)$, where $g$ is a function with changes in the first derivative penalized (\ie changes away from linear are penalized).
The $\logit$ link function both limits $\pi_{i,t}$ to the interval $[0, 1]$, and logit-linear growth approximates exponential growth for small $\pi_{i,t}$~\autocite{ealesAppropriately}.
Combined with property 4, the form of $g$ the null model is $g(i, t) = a_0 + a_t t + f(i)$ where $a_0$ and $a_t$ are model parameters and $f$ is a function to be specified; that is, changes away from this model should be penalized.

>>>>

First, in \cref{backcalc}, I use a backcalculation approach, based on \cref{E-inc-prev:eq:EPt}.
This approach enables estimation of incidence within (sub)populations of England.
There are several attractive features about this approach in the context here.
The backcalculation approach makes no assumptions about how transmission occurs; the assumptions have the potential to introduce bias.
It can also easily incorporate arbitrary covariates to adjust for non-response bias.
The scale of the CIS and desire to correct for non-response bias presents challenges for this methodology, especially within the limited computation of the SRS environment.

@@

% \subsubsection{Possible approaches}

The first approach, which I call the \emph{incidence-inferred} approach, is a single statistical model where incidence is inferred, with prevalence as a derived quantity.
The basis of the statistical model, in a Bayesian framework, would be as follows.
\begin{align}
    \vec{Z} \mid \vec{\phi} &\dist g_Z(\vec{\phi}) \\
    \vec{P_{i}} &= \matr{S} \vec{Z}_i \\
    Y_{i,t} \mid \vec{P} &\dist \BB(n_{i,t}, P_{i,t})
\end{align}
where $\vec{Z}$ is a vector of incidence proportion for each strata and day defined by some model $g$ with parameters $\vec{\phi}$; $P_{i,t}$ is the prevalence in stratum $i$ on day $t$; $\matr{S}$ is the survival matrix defined in \cref{E-inc-prev:eq:Smatrix}; $Y_{i,t}$ is the number of positive tests in stratum $i$ on day $t$; and $n_{i,t}$ is the number of tests in stratum $i$ on day $t$.
The model $g_Z$ would be specified to share information across time and/or strata.
This both helps stabilise the estimation (see \cref{intro:sec:disease-model}) and allows borrowing of information across strata.
Borrowing information is required because some strata and time combinations have few positives (fewer than five per day).
The relationship between $P$ and $Z$ is from \cref{E-inc-prev:eq:EPt-matrix}, although dividing both sides by $\Npop$ to have the values in terms of incidence proportion and prevalence rather than incidence and number of prevalent individuals.

The second approach, which I call the \emph{prevalence-inferred} approach takes a two-step approach.
First, it uses the statistical model from \cref{E-biology-data:sec:MRP} to estimate prevalence.
Then, it transforms the posterior distribution of prevalence to a posterior distribution of incidence using the forward substitution as in \cref{E-inc-prev:eq:forward-substitute}.
At a high-level, the statistical model is:
\begin{align}
    \vec{P} \mid \vec{\phi} &\dist g_P(\vec{\phi}) \\
    Y_{i,t} \mid \vec{P} &\dist \BB(n_{i,t}, P_{i,t})
\end{align}
where $g_P$ is analogous to $g_Z$ in providing a joint prior for the prevalence in each stratum and day combination.
Step 2 is...

\subsection{Desirable properties}

The incidence-inferred approach is more parsimonious in being a single step procedure.
However, it poses several pragmatic challenges which means I favour the prevalence-inferred approach.
Here I explain that further developments, beyond the scope of this thesis, would be required to use the incidence-inferred approach with a suitable $g_Z$ within the computational constraints of the SRS.

@@

The chosen approach should have the following properties.
\begin{enumerate}
    \item The smoothing model, $g_Z$ or $g_P$, over time should penalise changes from continuous exponential growth of $\pi$ over time.
        Penalizing changes from continuous exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalization has previously been shown to improve estimates of SARS-CoV-2 prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}. Used by CIS (section 3.?).
    \item Limit estimates to epidemiologically and biologically plausible values.
        That is, incidence proportion and prevalence should be in the interval $[0, 1]$.
        % While incidence proportion can theoretically be greater than 1, this would imply that some individuals were infected multiple times in a day which is not biologically plausible.
    \item Have sufficient parameters to incorporate all necessary covariates to correct for non-response bias.
    \item Borrow information across the different covariates.
        The borrowing should penalize changes in the distribution of incidence/prevalence across strata in the same geographic region being constant.
        A constant distribution is predicted by the theoretical study of epidemics, for example the dynamics of stratified, mechanistic transmission models (such as the one introduced in \cref{E-SEIR:sec:structured-populations}) over short periods of time with negligible behavioural and immunity changes.
    \item Be computationally efficient.
        The model will be fit to granular data and therefore needs to be run within the SRS, with its computational limitations (see \cref{E-biology-data:sec:SRS}).
\end{enumerate}

Properties 1 and 2 suggest a model of the form $\logit(\pi_{i,t}) = g(i, t)$, where $g$ is a function with changes in the first derivative penalized w.r.t.\ $t$ (\ie changes away from linear are penalised).
The $\logit$ link function both limits $\pi_{i,t}$ to the interval $[0, 1]$, and logit-linear growth approximates exponential growth for small $\pi_{i,t}$~\autocite{ealesAppropriately}.
Combined with property 4, the form of $g$ the null model is $g(i, t) = a_0 + a_t t + f(i)$ where $a_0$ and $a_t$ are model parameters and $f$ is a function to be specified; that is, changes away from this model should be penalized.
